A Linear Solution of the Linearity Challenge

### 2025-12-23 paper/calculus.tex
summary of section 2

The calculus of pure sessions that we formalize is dubbed \Calculus and is a
linear version of \CP~\citep{Wadler14,GayVasconcelos25} such that every session
endpoint is meant to be used \emph{exactly once}. This differs from \CP and
other session calculi where session endpoints can be used repeatedly, albeit
sequentially. The key idea is that each message sent on a session endpoint is
paired with a \emph{fresh continuation channel} on which the rest of the
conversation takes place.
%
This way of encoding sessions using explicit continuations has been first
described by \citet[extended version]{Kobayashi02b} and later studied by
\citet{DardhaGiachinoSangiorgi17}.

STRUCTURE OF SECTION

>>>>

As we have anticipated in \Cref{sec:introduction}, \LCP is a linear version of
\CP~\citep{Wadler14,GayVasconcelos25} such that every linear channel is meant to
be used for \emph{exactly one} communication. This differs from \CP and other
session calculi where session endpoints can be used repeatedly, albeit
sequentially. A session consisting of several messages exchanged in sequence can
be encoded using linear channels by pairing each message with a \emph{fresh
continuation channel} on which the rest of the session takes place.
%
This way of encoding sessions using explicit continuations has been first
described by \citet[extended version]{Kobayashi02b} and later studied by
\citet{DardhaGiachinoSangiorgi17}.

In this section we present \LCP, starting from its types (\Cref{sec:types}),
which coincide with classical linear logic propositions, its syntax
(\Cref{sec:syntax}), the reduction semantics (\Cref{sec:semantics}), the typing
rules (\Cref{sec:typing-rules}) and the formulation of the properties ensured by
the type system (\Cref{sec:properties}).
